
Please do not worry about my chapter 2 missing from here; I did the earleir exercises with my home PC but this week I am working with my laptop and there is a problem knitting chapter 2 from my laptop because the file directories I've used there are for my home PC...

**Here begins Week 3**

*Data Wrangling*

The Data Wrangling was done in the file create_students.R

*Logistic Regression*

Using alc_data

The data has 35 variables and 382 observations. I am picking the following variables to test in the logistic regression model: sex, age, health and romantic. The hypothesis is that males will be more likely than females to consume high quantities of alcohol, older people will be more likely than younger people to consume high quantities of alcohol, people with more absences are going to use more alchohol, and people who are not in a romantic relationship are more likely to consume higher quantities of alcohol than people in romantic relationships.

Visualizing data


```{r qplot, fig.width=4, fig.height=3, message=FALSE}
alc_data <- read.csv("Z:\\alc_data", header=TRUE)
library(ggplot2)

qplot(alc_use,data=alc_data)
qplot(sex,data=alc_data)
qplot(age,data=alc_data)
qplot(health,data=alc_data)
qplot(romantic,data=alc_data)

g2 <- ggplot(alc_data, aes(high_use))
g2 + facet_wrap("sex") + geom_bar()

g2 <- ggplot(alc_data, aes(x = high_use, y = age))
g2 + geom_boxplot() + ylab("age")+ggtitle("high alcohol use ang age")
g3 <- ggplot(alc_data, aes(x=high_use, y=absences))
g3 + geom_boxplot() + ylab("absences")+ggtitle("high alcohol use ang absences")

table(alc_data$high_use,alc_data$sex)
table(alc_data$high_use,alc_data$romantic)
```


```{r, building the model}
alc_data <- read.csv("Z:\\alc_data", header=TRUE)
m <- glm(high_use ~ sex + age + romantic + absences, data = alc_data, family = "binomial")
library(dplyr)
print(m)
summary(m)
coef(m)
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp
cbind(OR, CI)
```

First from the summary of the model we see that sex and absences are significant variables, but romantic and age are not. Thus we shall only interpret the coefficients for sex and absences.

We turn to the ORs for intepretation and find that the odds for men to have high alcohol use are 2.7 [1.7, 4.4] times greater than for females. With absences, for every one-unit increase on the absence scale, the odds of high alcohol use are 1.07 [1.04, 1.11] greater, which may at first seem like a small difference but we need to remember that absences is a contintuous variable and one unit increase is a small unit. In sum, sex and absences are significant predictors for high alcohol use. The hypotheses for these variables were correct (i.e. males hae greater odds of using high alc, and people with more absences are more likely to use high alc as well) but the hypotheses for age and romantic were wrong since these are not significant predictors in the mode.

Next I added the probabilities to the data and for the selected variables took a closer look and then created a table for the predictions and explored the predictions
```{r, probabilities}
alc_data <- read.csv("Z:\\alc_data", header=TRUE)
probabilities <- predict(m, type = "response")
alc_data <- mutate(alc_data, probability = probabilities)
alc_data <- mutate(alc_data, prediction = probability > 0.5)
select(alc_data, absences, sex, high_use, probability, prediction) %>% tail(10)
table(high_use = alc_data$high_use, prediction = alc_data$prediction)
library(ggplot2)
g <- ggplot(alc_data, aes(x = probability, y = high_use, col = prediction))
g + geom_point()

table(high_use = alc_data$high_use, prediction = alc_data$prediction)
```

I tried to do the next extra step too but I get a NaN result which probably isn't good...
```{r, probabilities2}
alc_data <- read.csv("Z:\\alc_data", header=TRUE)

loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}
loss_func(class = alc_data$high_use, prob = alc_data$probability)
```
