# Week 2 exercises

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

#First part, Data wrangling

The data wrangling was done in the file create_learning2014.R with the following code, comments in *italics*


*creating questions*
```deep_questions <-c("D03", "D11", "D19", "D27", "D07", "D14", "D22", "D30","D06",  "D15", "D23", "D31")```
```surface_questions <- c("SU02","SU10","SU18","SU26", "SU05","SU13","SU21","SU29","SU08","SU16","SU24","SU32")```
```strategic_questions <- c("ST01","ST09","ST17","ST25","ST04","ST12","ST20","ST28")```

*creating columns*


```deep_columns <- select(JYTOPKYS3_data one_of(deep_questions))
JYTOPKYS3_data$deep <- rowMeans(deep_columns)```

```surface_columns <- select(JYTOPKYS3_data, one_of(surface_questions))
JYTOPKYS3_data$surf <- rowMeans(surface_columns)```

```strategic_columns <- select(JYTOPKYS3_data, one_of(strategic_questions))
JYTOPKYS3_data$stra <- rowMeans(strategic_columns)```

*assigning which columns to keep*
```keep_columns <- c("gender", "Age", "Attitude", "deep", "stra", "surf", "Points")
learning2014 <- select(JYTOPKYS3_data, one_of(keep_columns))```

*filtering points*
```learning2014 <- filter (learning2014, Points >0)```

*the data now has  166 observations and 7 variables*
*setting working directory from the menu and then saving*
```write.csv(learning2014, file="learning2014", append = TRUE, quote = TRUE, sep = " ",
          eol = "\n", na = "NA", dec = ",", row.names = TRUE,
          col.names = TRUE, qmethod = c("escape", "double"),
          fileEncoding = "")```
```?write.csv```
```read.csv("learning2014")```
```str(learning2014)```
```head(learning2014)```

#Second part, using the data & regression analysis
*Step 1: opening the data and exploring*
I opened the data by using the menu, I notice that there is an extra variable X1 for the first variable which seems to be a numerical count of the rows... I deleted it with

```learning2014 <- select(learning2014, -X1)```

The learning2014 dataset has 166 observations (no missing values), and 7 variables: "gender"   "Age"      "Attitude" "deep"     "stra"     "surf"     "Points"  (accessed with command 'names')
I also explored the data with the following commands
```
library(dplyr)
dim(learning2014)
names(learning2014)```

*Step 2 show a graphical overview of the data and show summaries of the variables in the data*

SOMETHING IS WRONG HERE, can't knit to html...
```{r qplot, fig.width=4, fig.height=3, message=FALSE}
library(ggplot2)
summary(learning2014)
qplot(Age,data=learning2014)
```

```{r qplot1, fig.width=4, fig.height=3, message=FALSE}
library(ggplot2)
qplot(gender,data=learning2014)
```
```{r qplot2, fig.width=4, fig.height=3, message=FALSE}
library(ggplot2)
qplot(Attitude,Age,data=learning2014)
```
